### 2013-10-10 paper/paper.tex

%-----------------------------------------------------------------------------
%
%			Edwin's and Franck's Paper !
%  uses the template for sigplanconf LaTeX Class
%
%-----------------------------------------------------------------------------

\documentclass{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\newcommand{\cL}{{\cal L}}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{PLPV ’14, Co-located with POPL 2014}{January 21, 2014, San Diego, CA, USA.}
\copyrightyear{2013} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\preprintfooter{Our paper for PLPV’14}
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Proof by reflection for equalities in algebraic structures}
%\subtitle{An example for a collection of tactics solving equalities in algebraic structures}

\authorinfo{Edwin Brady\and Franck Slama}
           {University of St Andrews, United-Kingdom}
           {ecb10@st-andrews.ac.uk \and fs39@st-andrews.ac.uk}
\bibliographystyle{abbrvnat}

\maketitle

\nocite{*}

\input{./content/abstract.tex}

\input{./content/introduction.tex}

\input{./content/idrisOverview}

\section{Normalization and polynomials in normal form}
% Ici on donne l'idée générale. Nom de la section peut etre à revoir
Lots of text.

More text.

\section{Datatypes}

Lots of text.

\subsection{Hierarchy of algebraic theorys}

Lots of text.

The exemple with Nat is a monoid

More text.

\subsection{Reflection in an abstract syntax}

Lots of text.

More text.

\section{Normalization procedures}

\subsection{Normalization for terms in a magma}

\subsection{Normalization for terms in a semi-group}

\subsection{Normalization for terms in a monoid}

\subsection{Normalization for terms in a group}

\subsection{Normalization for terms in a ring}

\subsection{Normalization for terms in a commutative ring}

\section{Coating}
%enrobage

\section{Metaification}

\section{Conclusion}

\acks

Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliography{biblio}

\appendix
\section{Appendix Title}

This is the text of the appendix, if you need one.

\end{document}

%                       Revision History
%                       -------- -------
%  Date         Person  Ver.    Change
%  ----         ------  ----    ------

%  2013.06.29   TU      0.1--4  comments on permission/copyright notices