### 2015-09-30 NotesOnCorrectnessAndCompleteness/notesOnCorrectnessAndCompleteness.tex
added various files lying around

% Franck's notes on correctness and completeness !
% last revision : 30 Sept 2015
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage[literate]{idrislang} 
%\usepackage{multicol}

%
\begin{document}
%
\frontmatter          % for the preliminaries
%
\pagestyle{headings}  % switches on printing of running heads
%\addtocmark{Hamiltonian Mechanics} % additional mark in the TOC
%

\mainmatter              % start of the contributions
%
\title{Notes on correctness and completeness of the approach we've followed and comparison with Coq} 
%
\titlerunning{Notes on correctness and completeness}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Franck Slama}
%
\authorrunning{Franck Slama} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%%%\tocauthor{}
%
\institute{University of St Andrews, St Andrews, United Kingdom\\
\email{fs39@st-andrews.ac.uk}
%,\\ WWW home page:
%\texttt{http://users/\homedir iekeland/web/welcome.html}
}

\maketitle              % typeset the title of the contribution

Our correctness lemma :
$\forall x y,\ norm (T(x)) = T(y) \rightarrow x = y$

If the normalisation of the reflection of x gives the reflection of y, then x=y

\end{document}