### 2016-02-16 ProofsAndTests/paper-TAP/content/4-testingInside.tex
Structure of the paper

\section{Testing the predicates inside the proof assistant}

- it is often easier to generate things rather than to characterise them 
[a quarter of a page : from 3.00 to 3.15]

- testing the structural properties by automatic generation of the terms that should satisfy them, and running a decision procedure on them.
[half a page : from 3.15 to 3.45]

- Going a step further : ``test automation by proof", which automatises all of them at once, by doing (another) formal proof, which shows that there is equivalence between all the terms generated and the original predicate that we wanted to certify correct according to the original specification.
[half a page : from 3.45 to 4.15]