### 2016-02-16 ProofsAndTests/paper-TAP/content/5-conclusion.tex

\section{Conclusions, related work, and future work (a quarter of a page)}
[from 4.15 to 4.30]