### 2016-02-19 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex

\section{``Structural" VS ``on content" properties}

\label{sect:differentKindOfProperties}

- differences between ``structural" properties, and properties about the ``content".
[half a page : from 2.30 to 3.00]