### 2016-02-20 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
paper almost finished

\section{Replacing tests on the predicate by proofs}

\label{sect:aStepForward}

- Going a step further : ``test automation by proof", which automatises all of them at once, by doing (another) formal proof, which shows that there is equivalence between all the terms generated and the original predicate that we wanted to certify correct according to the original specification.
[half a page : from 3.45 to 4.15]