### 2016-04-28 ProofsAndTests/paper-TAP/content/4-conclusion.tex
added the file with the updated name

\section{Conclusions and future work}
\label{sect:Conclusion}

In this paper we've presented a new way to test a predicate based on an automatic generation of terms that should have the desired property. This adequacy between the predicate and the generator helps to gain confidence in the predicate. The technique presented on section \ref{sect:testingInside} was based on a finite observation of the terms generated in endless amount, processed by the decision procedure. We have also shown on an example how these tests can be automated.

We haven't been able to find much work done in the direction of predicate testing in the environment of proof assistants, but we strongly believe that this aspect is crucial, as there is absolutely no point to prove the ``correctness" of a function relatively to a bad specification. One could however question why formal certification is needed at all, if after going through all the effort of interactive theorem proving we still have to test the specification itself, and also need tools to support it. We believe that the process of formalisation helps to uncover things that were missing or weakly specified in the informal requirements. For example, the development of the formally verified compiler CompCert~\cite{Krebbers2014} has contributed to brought to light many under-specified behaviours in the specification of the C standard.

The machinery developed for the running example presented in this paper is extremely specific and it is not reasonable to believe that this work should and could be done for every formal specification. What it shows is that we really need to explore how proof assistants themselves could help to gain confidence about predicates and logical formulae. A possible direction could be to build execution engines for formal specifications written in dependent type theories. Such a system would take in input a predicate and would produce some of the terms that make this predicate hold. The ideal would be to have a query system where one could ask the system to try to look if some specific terms are captured by the predicate. Since the problem of finding proof is undecidable in the general case in higher-order logics (that's also the case in first-order logic), such a system can't be complete and entirely automatic, and therefore the user would have to help the system at times.

Another possible direction is to equip proof assistants with many robust and generic concepts (like being sorted) once and for all. That would save the user from the error prone activity of writing many primitive logical properties. Equipping proof assistants with many generic and useful concepts already available, like bricks ready to be assembled, is another current challenge to make proofs assistants really usable.