### 2017-05-16 paper-CICM-17/paper.tex

% Franck Slama's and Edwin Brady's paper for CICM-2017
% That follows the LLNCS style from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{float}
\usepackage{url}
\usepackage{listings}

\usepackage{tikz}
\usetikzlibrary{matrix}

\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

\def\code#1{{\mbox{\texttt{#1}}}} % For short element of code in the test that should not be hyphenated
\def\codeL#1{{\texttt{#1}}} % For Long element of code in the text that might need to be hyphenated

\newcommand{\figrule}{\begin{center}\hrule\end{center}}

%
\begin{document}
%
\pagestyle{headings}  % switches on printing of running heads

\mainmatter              % start of the contributions
%
\title{Automatically Proving Equivalence by Type-Safe Reflection}
%\subtitle{Or how to use dependent types to address some of the problems they have caused}
%
\titlerunning{xxx}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Franck Slama \and Edwin Brady}
\institute{University of St Andrews, Scotland, UK\\
\email{fs39@st-andrews.ac.uk, ecb10@st-andrews.ac.uk}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
One difficulty with programming with dependent types is that proof obligations
arise naturally once programs become even moderately sized. For example,
implementing an adder for binary numbers indexed over their natural number
equivalents naturally leads to proof obligations for equalities of expressions
over natural numbers. The need for these equality proofs comes, in intensional
type theories, from the fact that the propositional equality enables us to
prove as equal terms that are not judgementally equal, which means that the
typechecker can't always obtain equalities by reduction.  As far as possible,
we would like to solve such proof obligations automatically. In this paper, we
show one way to automate these proofs by reflection in the dependently typed
programming language Idris.  We show how defining reflected terms indexed by
the original Idris expression allows us to construct and manipulate proofs. We
build a hierarchy of tactics for proving equivalences in semi-groups, monoids,
commutative monoids, groups, commutative groups, semi-rings and rings. We also
show how each tactic reuses those from simpler structures, thus avoiding
duplication of code and proofs.

\keywords{proof automation, equivalence, equality, proof by reflection, correct-by-construction software, type-driven development}
\end{abstract}
%

\input{./content/1-introduction.tex}

% The chap 2 on the smaller tactic for Nat has been removed for space constraint

\input{./content/3-hierarchyOfTactics.tex}

\input{./content/4-relatedWork.tex}

\input{./content/5-conclusions.tex}

\bibliographystyle{splncs03}
\bibliography{dtp}

\end{document}