### 2017-05-20 paper-CICM-17/content/5-conclusions.tex

\section{Results, conclusions and future work}

We have implemented a very generic solution to the initial problem of index
mismatch that appeared in section~\ref{sect:motivatingExample} when we were
using indexed types. This solution takes the form of a hierarchy of provers for
equivalences in algebraic structures. These provers are generic in multiple
ways: they work for many algebraic structures (semi-group, monoid, commutative
monoid, group, commutative group, ring and semi-ring), for any type that
behaves as one of these structures, and for any equivalence relation on this
type (and not only for the propositional equality). The implementation is
modular and each prover reuses the prover of the structure from which it
inherits. These provers can automatically prove equivalences between terms,
which discharges the user of providing proofs for some proof obligations, as
\code{adc\_lemma\_2}. Thus, these provers enable the user to focus on the
interesting proofs that requires specific knowledge and creativity, instead of
wasting time for proving routine lemmas that can be automated.

The correct-by-construction method that we have followed has involved the
design of a type-safe reflection mechanism where reflected terms are indexed
over the concrete inputs, and from which we are able to pull out the proofs
easily. The construction of the proof is done step by step, and it follows the
construction of the normalised terms. Therefore, unlike Coq's and Agda's ring
provers, we do not have the traditional duplication that happens when
separating the computational content from the proof of correctness. Instead, in
our approach, the construction of the proof is guided by the construction of
the normalised terms, which, in addition of avoiding redundancy, simplifies the
proof generation considerably. This development shows that if dependent types
effectively bring some new problems, they are also very expressive tools for
building correct-by-construction software where the development is \emph{driven
by the types}.

This work can be extended to build new provers for less common algebraic
structures and for more specific structures. For example, regular expressions
form a ``pre semi-ring" with some extra axioms. We will refactor the semi-ring
level with the creation of the intermediate structure of pre semi-ring, that
will not necessary have the property that $0$ is an annihilator element for the
product. Then, we could build a specific prover for regular expressions, that
would use the normalisation function of the pre semi-ring level and that would
only have to deal with the specific properties of regular expressions : the
neutral element $\varnothing$ for the concatenation of languages is also a
neutral element for the product of languages, the idempotence of the addition
of languages, and the rules of simplifications for the new Kleene star
operation.