### 2017-05-20 paper-CICM-17/content/5-conclusions.tex

\section{Results, conclusions and future work}

We have implemented a generic solution to the initial problem of index
mismatch (section~\ref{sect:motivatingExample}) when 
using indexed types. This solution takes the form of a hierarchy of provers for
equivalences in algebraic structures. These provers are generic in several
ways: they work for many algebraic structures (semi-group, monoid, commutative
monoid, group, commutative group, ring and semi-ring); for any type that
behaves as one of these structures; and, for any equivalence relation on this
type. The implementation is
modular and each prover reuses the prover of the structure from which it
inherits. These provers can automatically prove equivalences between terms,
so the user need not prove obligations by hand,
like \code{adc\_lemma\_2}. Thus, these provers enable the user to focus on the
interesting proofs that requires specific knowledge and creativity, instead of
routine, automatable, lemmas.

Our correct-by-construction method involved the design of a type-safe
reflection mechanism where reflected terms are indexed over concrete inputs,
and from which we are able to extract proofs directly.  Unlike Coq's and Agda's
ring provers, we do not have the duplication that arise when separating the
computational content from the proof of correctness. Instead, construction of
the proof is done step by step, following the construction of the normalised
terms.  In addition to avoiding redundancy, this simplifies the proof
generation considerably. This development shows that if dependent types
effectively bring some new problems, they are also very expressive tools for
building correct-by-construction software where the development is \emph{driven
by the types}.

This work can be extended to build new provers for less common algebraic
structures and for more specific structures. For example, regular expressions
form a ``pre semi-ring'' with some extra axioms. We will refactor the semi-ring
level with the creation of the intermediate structure of pre semi-ring, that
will not necessary have the property that $0$ is an annihilator element for the
product. Then, we could build a specific prover for regular expressions, that
would use the normalisation function of the pre semi-ring level and that would
only have to deal with the specific properties of regular expressions : the
neutral element $\varnothing$ for the concatenation of languages is also a
neutral element for the product of languages, the idempotence of the addition
of languages, and the rules of simplifications for the new Kleene star
operation.